---
title: "Untitled"
output: github_document
---

# Importing Tracab Tracking Data into R

Firstly, chose a match you want to import and put the .dat file into the working directory of R. Let's read it to a dataframe. 

```{r}
# change below to the name of the tracking data file
match <- "TrackingDataForAMatch.dat"
# use the read.table function to easily import the data 
dat <- read.table(tc)
```

The data is now read into R, but... it is not organised and spilt into a useable structure. We want to have a new row of data for every player for each frame (25 times a second). Let's setup a dataframe to store this information in. 

```{r}
seg <- data.frame  (team_HA=character(),
                   TrackID=character(),
                   JerseyNo=character(),
                   x=character(),
                   y=character(),
                   Speed=character(),
                   stringsAsFactors=FALSE)
```

In the dat object we currently have one row for every frame, in that row there is one massive string which contains the positional data of each player for that frame. We want to iterate through the dataset and break each row up. 

```{r}
for (i in dat$V1[1:nrow(dat)]){
frame <- unlist(strsplit(i, ":"))[1]
x2 <- unlist(strsplit(i, ":"))[2] 
x3 <- unlist(strsplit(x2,";"))
tempX <- data.frame(matrix(unlist(strsplit(x3,",")), nrow=29, byrow=T))
tempX$frame <- frame
colnames(tempX) <- c("team_HA", "TrackID", "JerseyNo", "x","y","Speed","frame")

x4 <- unlist(strsplit(i, ":"))[3]
tempX2 <- data.frame(matrix(unlist(strsplit(x4,",")), nrow=1, byrow=T))

## create the referee data 
team_HA = "10"
TrackID = "30"
JerseyNo = NA
x = tempX2$X1
y = tempX2$X2
Speed = tempX2$X3

# create the ball data 
ballTemp <- data.frame(team_HA,TrackID,JerseyNo,x,y,Speed)
ballTemp$frame <- frame
catch <- rbind(tempX,ballTemp)
catch <- rbind(catch[which(catch$team_HA==1),] , (catch[which(catch$team_HA==0),]), (catch[which(catch$team_HA==10),]))

# join the row data to the overall seg dataframe
seg <- rbind(seg,catch)
cat(".")
}
```
There is still some cleaning up of the data to do 

```{r}
# clean up the seg data frame 
seg$TrackID[is.na(seg$TrackID)] <- "50" # set the ball trackID
seg$x <- as.numeric(as.character(seg$x)) # convert from a string to a numeric
seg$y <- as.numeric(as.character(seg$y)) # convert from a string to a numeric
seg$frame <- as.numeric(as.character(seg$frame)) # convert from a string to a numeric
```

Now we have the tracking data in a structure that is more flexible and ready to utilise in many ways. 
